A Logical Solution of the Linearity Challenge

### 2025-11-29 paper/Context.lagda.tex
included reduction
restroed contexts and permutations

\begin{code}[hide]
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)

@@

We are going to adopt a nameless representation of channels whereby a channel is
identified by its position in a typing context. This representation is akin to
using De Bruijn indices, except that the index, instead of being represented
explicitly by a natural number, is computable from the \emph{proof} that the
type of the channel belong to the typing context.
%
Typing contexts are represented using \emph{lists} of types, where the
(polymorphic) \AgdaDatatype{List} type is defined in Agda′s standard library.

@@

+-unit-r = +-comm +-unit-l
\end{code}

@@

+-sing-l : ∀{A B Γ} → [ A ] ≃ B , Γ → A ≡ B × Γ ≡ []
+-sing-l (split-l split-e) = refl , refl

data Un : Context → Set where
  un-[]  : Un []
  un-∷   : ∀{A Γ} → Un Γ → Un (¿ A ∷ Γ)
\end{code}

>>>>

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_; map)

@@

We are going to adopt a nameless representation of channels whereby a channel is
identified by its position in a typing context. This representation is akin to
using De Bruijn indices, except that the index, instead of being represented
explicitly by a natural number, is computable from the \emph{proof} that the
type of the channel belong to the typing context.
%
Typing contexts are represented using \emph{lists} of types, where the
(polymorphic) \AgdaDatatype{List} type is defined in Agda's standard library.

@@

+-unit-r = +-comm +-unit-l

++≃+ : ∀{Γ Δ} → Γ ++ Δ ≃ Γ + Δ
++≃+ {[]} = +-unit-l
++≃+ {_ ∷ _} = split-l ++≃+
\end{code}

@@

+-sing-l : ∀{A B Γ} → [ A ] ≃ B , Γ → A ≡ B × Γ ≡ []
+-sing-l (split-l split-e) = refl , refl
\end{code}

To conclude, we define a predicate $\AgdaDatatype{Un}~\Context$ that holds
whenever $\Context$ is \emph{unrestricted}, namely when it is made of types of
the form $?A$. We will need this predicate in the definition of a server, which
must obey the typing rule \ServerRule of \Cref{tab:typing-rules}.

\begin{code}
data Un : Context → Set where
  un-[]  : Un []
  un-∷   : ∀{A Γ} → Un Γ → Un (`? A ∷ Γ)
\end{code}

### 2025-11-29 paper/Permutations.lagda.tex
restroed contexts and permutations

\begin{code}[hide]
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)

@@

Just like permutations from Agda's standard library, also \AgdaDatatype{\_\#\_}
enjoys a number of expected (and useful) properties, such as symmetry. Among all
properties, in the following we will make use of a specific one that connects
context splitting and permutations: if $\Context \simeq A, \ContextD$, then
$\Context$ must be a permutation of $A~\AgdaInductiveConstructor{∷}~\ContextD$.

@@

#shift : ∀{Γ A Δ} → (Γ ++ A ∷ Δ) # (A ∷ Γ ++ Δ)
#shift {[]} = #refl
#shift {B ∷ Γ} = #tran (#next #shift) #here

>>>>

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)

@@

Just like permutations from Agda's standard library, also \AgdaDatatype{\_\#\_}
enjoys a number of expected (and useful) properties, such as symmetry. Among all
properties, in the following we will make use of a specific one that connects
context splitting and permutations: if $\Context \simeq A, \ContextD$, then
$\Context$ must be a permutation of
$A~\AgdaInductiveConstructor{∷}~\ContextD$.\Luca{Ci sono molte altre proprietà
non descritte, controllare quelle che servono davvero.}

@@

#shift : ∀{A Γ Δ} → (Γ ++ A ∷ Δ) # (A ∷ Γ ++ Δ)
#shift {_} {[]} = #refl
#shift {_} {B ∷ Γ} = #tran (#next #shift) #here
